Supporting information: Copulas and their potential for ecology

### 2019-09-22 Paper.Rmd
edits missed before

We present green spruce aphid abundance (Table \ref{tab_data_info}) results
as an example multivariate analysis.
Independence was rejected for each pair of the 10 sampling locations. 
Best-fitting (lowest-AIC) copulas were non-N
for the large majority of location pairs (Table \ref{SM-tab_count_bestcop}), and 
AIC weights for best-fitting copulas (Table \ref{SM-tab_count_AICw}) 
were typically substantially 
larger than normal copula weights (Table \ref{SM-tab_count_AICw_normal}), answering the first part of Q1. Goodness of fit tests 
(Tables \ref{SM-tab_count_p_CvM}, \ref{SM-tab_count_p_KS}) in every case 
failed to reject the hypothesis that the AIC-best copula family was also 
an objectively adequate description of 
the data; i.e., the collection of 
copula families we used was sufficiently broad to characterize these data.
Model-averaged lower- (Table \ref{SM-tab_count_LTdep_AICw}) and upper-tail
(Table \ref{SM-tab_count_UTdep_AICw}) dependence statistics were exclusively 
positive and were frequently large, thereby differing 
from what a normal copula would give (i.e., 0), and answering the second
part of Q1: these data have greater tail dependence 
(lower and upper) than would be expected from an N-copula null hypothesis. 
Model-averaged lower- minus upper-tail dependence statistics 
(Table \ref{SM-tab_count_LTmUT}) were positive 
for all but a few location pairs (`r sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)` 
out of `r (dim(RES_aphid_count$gfc_p_CvM)[1]^2-dim(RES_aphid_count$gfc_p_CvM)[1])-2*RES_aphid_count$num_indep_loc_pair`). Thus the spatial synchrony of rarity in the green spruce aphid is stronger 
than the spatial synchrony of outbreaks. This also answers the third part of Q1.
Nonparametric statistics verified that tail dependence was asymmetric for the green 
spruce aphid abundance data (Appendix \ref{SM-Q1multvarnonparam}, \ref{SM-spatial_resamp}).

@@

<!--Displaying a table with npa summary results for soilCN data-->
```{r tab_soilCN_asym,echo=F,results="asis"}
tab_biv_npa_summary<-readRDS("./Results/stat_results/tab_biv_npa_summary.RDS")
tab_soilCN_asym<-tab_biv_npa_summary[1:3,2:4]
knitr::kable(tab_soilCN_asym, booktabs = T, escape=F,
             format = "latex",linesep = "",
      caption.short = "Nonparametric asymmetry tests against an N-copula null, soil C and N",
      caption = "Soil C and N dataset results for nonparametric tests of whether asymmetry of tail dependence was significant, compared to a normal copula null hypothesis. The values of the listed statistics (cor$_{FS}$-cor$_{LS}$ = $\\cor_{0,0.1}-\\cor_{0.9,1}$, P$_{FS}$-P$_{LS}$ = $\\Ps_{0,0.1}-\\Ps_{0.9,1}$, D$_{LS}^2$-D$_{FS}^2$ = $\\Dsq_{0.9,1}-\\Dsq_{0,0.1}$; here FS stands for `first slice' and LS for `last slice') for real data were compared to their values for 1000 Kendall- or Spearman-preserving normal surrogates, in seperate comparisons. A table entry $<X$ indicates the value of the given statistic on the data was less than its value on $X$ of the surrogates, so entries of the form $<X$ for $X$ equal to $975$ or above indicate that upper-tail dependence was significantly stronger than lower-tail dependence.\\label{tab_soilCN_asym}")
```

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/stat_results/stat_soilCN/soilCN_bivfunctionplot.pdf}
\caption[Nonparametric tests against an N-copula null, soil C and N data]{Nonparametric tests for tail dependence and other deviations from normal copula structure, for soil C and N data. As described in the main text, the values of the statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$ and $\Dsq_{l_b,u_b}$ for real data (red crosses) were compared to distributions of their values on 1000 Kendall- or Spearman-preserving normal surrogates of the data (blue and green x's show 0.025 and 0.975 quantiles), separately in two comparisons for each of the ranges $(l_b,u_b)=(0,0.1),\ldots,(0.9,1)$. Whenever the red cross was outside the range given by the x's, text at the top of panels indicates the number of surrogate values the real-data value was greater than or less than. For instance, a value $>N$ (respectively, $<N$) means the value of the statistic on real data was greater than (respectively, less than) its value on $N$ surrogates. When the statistic was greater than 975 or less than 975 surrogate values, it indicates significance ($95\%$ confidence level). When $\cor$ or $\Ps$ values (respectively, $\Dsq$ values) were greater than surrogates, it means dependence in that part of the distributions was stronger than (respectively, weaker than) expected from a normal copula null hypothesis. \label{fig_soilCN_nonparam}}
\end{center}
\end{figure}

@@

\begin{figure}[!h]
\textbf{ \hspace{6.6 cm} (A) \hspace{4.6 cm} (B)} \\
\vspace{-0.4 cm}
\begin{center}
\includegraphics[width=5 cm]{./Results/skewness_results/skewness_aphid_count/skewness_aphid_count_sp_10_spearman.pdf}
\includegraphics[width=5 cm]{./Results/skewness_results/skewness_aphid_ff/skewness_aphid_ff_sp_11_spearman.pdf}\\
\end{center}
\vspace{0.2 cm}
\textbf{ \hspace{6.6 cm} (C) \hspace{4.6 cm} (D)} \\
\vspace{-0.4 cm}
\begin{center}
\includegraphics[width=5 cm]{./Results/skewness_results/skewness_plankton_north_sea/skewness_plankton_north_sea_sp_16_spearman.pdf}
\includegraphics[width=5 cm]{./Results/skewness_results/skewness_methane/skewness_methane_spearman.pdf}
\caption{Skewness of spatially averaged green spruce aphid abundance $(A)$, leaf-curling plum 
aphid first flight dates $(B)$, \emph{Ceratium furca} abundance $(C)$, and methane-flux
$(D)$ time series compared to a multivariate 
N-copula null hypothesis. Black dots are empirical 
skewnesses; see text 
and Appendix \ref{SM-Surrog_multi} for details of the null hypothesis. Results show a 
tendency for skewness of the spatial average to be affected as hypothesized 
by tail dependence. \label{fig_skewness_spearman}}
\end{center}
\end{figure}

>>>>

We present green spruce aphid abundance (Table \ref{tab_data_info}) results
as an example multivariate analysis.
Independence was rejected for each pair of the 10 sampling locations. 
Best-fitting (lowest-AIC) copulas were non-N
for the large majority of location pairs (Table \ref{SM-tab_count_bestcop}), and 
AIC weights for best-fitting copulas (Table \ref{SM-tab_count_AICw}) 
were typically substantially 
larger than normal copula weights (Table \ref{SM-tab_count_AICw_normal}), answering the first part of Q1. Goodness of fit tests 
(Tables \ref{SM-tab_count_p_CvM}, \ref{SM-tab_count_p_KS}) in every case 
failed to reject the hypothesis that the AIC-best copula family was also 
an objectively adequate description of 
the data; i.e., the collection of 
copula families we used was sufficiently broad to characterize these data.
Model-averaged lower- (Table \ref{SM-tab_count_LTdep_AICw}) and upper-tail
(Table \ref{SM-tab_count_UTdep_AICw}) dependence statistics were exclusively 
positive and were frequently large, thereby differing 
from what a normal copula would give (i.e., 0), and answering the second
part of Q1: these data have greater tail dependence 
(lower and upper) than would be expected from a normal-copula null hypothesis. 
Model-averaged lower- minus upper-tail dependence statistics 
(Table \ref{SM-tab_count_LTmUT}) were positive 
for all but a few location pairs (`r sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)` 
out of `r (dim(RES_aphid_count$gfc_p_CvM)[1]^2-dim(RES_aphid_count$gfc_p_CvM)[1])-2*RES_aphid_count$num_indep_loc_pair`). Thus the spatial synchrony of rarity in the green spruce aphid is stronger 
than the spatial synchrony of outbreaks. This also answers the third part of Q1.
Nonparametric statistics verified that tail dependence was asymmetric for the green 
spruce aphid abundance data (Appendix \ref{SM-Q1multvarnonparam}, \ref{SM-spatial_resamp}).

@@

<!--Displaying a table with npa summary results for soilCN data-->
```{r tab_soilCN_asym,echo=F,results="asis"}
tab_biv_npa_summary<-readRDS("./Results/stat_results/tab_biv_npa_summary.RDS")
tab_soilCN_asym<-tab_biv_npa_summary[1:3,2:4]
knitr::kable(tab_soilCN_asym, booktabs = T, escape=F,
             format = "latex",linesep = "",
      caption.short = "Nonparametric asymmetry tests against a normal-copula null, soil C and N",
      caption = "Soil C and N dataset results for nonparametric tests of whether asymmetry of tail dependence was significant, compared to a normal-copula null hypothesis. The values of the listed statistics (cor$_{FS}$-cor$_{LS}$ = $\\cor_{0,0.1}-\\cor_{0.9,1}$, P$_{FS}$-P$_{LS}$ = $\\Ps_{0,0.1}-\\Ps_{0.9,1}$, D$_{LS}^2$-D$_{FS}^2$ = $\\Dsq_{0.9,1}-\\Dsq_{0,0.1}$; here FS stands for `first slice' and LS for `last slice') for real data were compared to their values for 1000 Kendall- or Spearman-preserving normal surrogates, in seperate comparisons. A table entry $<X$ indicates the value of the given statistic on the data was less than its value on $X$ of the surrogates, so entries of the form $<X$ for $X$ equal to $975$ or above indicate that upper-tail dependence was significantly stronger than lower-tail dependence.\\label{tab_soilCN_asym}")
```

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/stat_results/stat_soilCN/soilCN_bivfunctionplot.pdf}
\caption[Nonparametric tests against a normal-copula null, soil C and N data]{Nonparametric tests for tail dependence and other deviations from normal copula structure, for soil C and N data. As described in the main text, the values of the statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$ and $\Dsq_{l_b,u_b}$ for real data (red crosses) were compared to distributions of their values on 1000 Kendall- or Spearman-preserving normal surrogates of the data (blue and green x's show 0.025 and 0.975 quantiles), separately in two comparisons for each of the ranges $(l_b,u_b)=(0,0.1),\ldots,(0.9,1)$. Whenever the red cross was outside the range given by the x's, text at the top of panels indicates the number of surrogate values the real-data value was greater than or less than. For instance, a value $>N$ (respectively, $<N$) means the value of the statistic on real data was greater than (respectively, less than) its value on $N$ surrogates. When the statistic was greater than 975 or less than 975 surrogate values, it indicates significance ($95\%$ confidence level). When $\cor$ or $\Ps$ values (respectively, $\Dsq$ values) were greater than surrogates, it means dependence in that part of the distributions was stronger than (respectively, weaker than) expected from a normal-copula null hypothesis. \label{fig_soilCN_nonparam}}
\end{center}
\end{figure}

@@

\begin{figure}[!h]
\textbf{ \hspace{6.6 cm} (A) \hspace{4.6 cm} (B)} \\
\vspace{-0.4 cm}
\begin{center}
\includegraphics[width=5 cm]{./Results/skewness_results/skewness_aphid_count/skewness_aphid_count_sp_10_spearman.pdf}
\includegraphics[width=5 cm]{./Results/skewness_results/skewness_aphid_ff/skewness_aphid_ff_sp_11_spearman.pdf}\\
\end{center}
\vspace{0.2 cm}
\textbf{ \hspace{6.6 cm} (C) \hspace{4.6 cm} (D)} \\
\vspace{-0.4 cm}
\begin{center}
\includegraphics[width=5 cm]{./Results/skewness_results/skewness_plankton_north_sea/skewness_plankton_north_sea_sp_16_spearman.pdf}
\includegraphics[width=5 cm]{./Results/skewness_results/skewness_methane/skewness_methane_spearman.pdf}
\caption{Skewness of spatially averaged green spruce aphid abundance $(A)$, leaf-curling plum 
aphid first flight dates $(B)$, \emph{Ceratium furca} abundance $(C)$, and methane-flux
$(D)$ time series compared to a multivariate 
normal-copula null hypothesis. Black dots are empirical 
skewnesses; see text 
and Appendix \ref{SM-Surrog_multi} for details of the null hypothesis. Results show a 
tendency for skewness of the spatial average to be affected as hypothesized 
by tail dependence. \label{fig_skewness_spearman}}
\end{center}
\end{figure}

### 2019-09-22 SupportingInformation.Rmd
edits missed before

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/BMR_results/birds/birds_numbin_10_bivfunctionplot.pdf}
\caption[Nonparametric tests against a normal copula null, bird body mass and BMR data]{Nonparametric tests for tail dependence and other deviations from normal copula structure, for bird body mass and BMR data. Same format as Fig. \ref{MT-fig_soilCN_nonparam}, see caption of that figure. \label{fig_bmr_birds_nonparam}}
\end{center}
\end{figure}

@@

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/BMR_results/mammals/mammals_numbin_10_bivfunctionplot.pdf}
\caption[Nonparametric tests against a normal copula null, mammal body mass and BMR data]{Nonparametric tests for tail dependence and other deviations from normal copula structure, for mammal body mass and BMR data. Same format as Fig. \ref{MT-fig_soilCN_nonparam}, see caption of that figure. \label{fig_bmr_mammals_nonparam}}
\end{center}
\end{figure}

@@

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/Ceder_creek_results/2000cc_numbin_4_bivfunctionplot.pdf}
\caption[Nonparametric tests against a normal copula null, Cedar Creek data]{Nonparametric tests for tail dependence and other deviations from normal copula structure, for Cedar Creek data, year 2000. As described in the main text, the values of the statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$ and $\Dsq_{l_b,u_b}$ for real data (red crosses) were compared to distributions of their values on 1000 Kendall- or Spearman-preserving normal surrogates of the data (blue and green x's show 0.025 and 0.975 quantiles), separately in two comparisons for each of the ranges $(l_b,u_b)=(0,0.25),\ldots,(0.75,1)$. If the red cross was outside the range given by the x's, text at the top of panels indicates the number of surrogate values the real-data value was greater than or less than. For instance, a value $>N$ (respectively, $<N$) means the value of the statistic on real data was greater than (respectively, less than) its value on $N$ surrogates. When the statistic was greater than 975 or less than 975 surrogate values, it indicates significance ($95\%$ confidence level). When $\cor$ or $\Ps$ values (respectively, $\Dsq$ values) were greater than surrogates, it means dependence in that part of the distributions was stronger than (respectively, weaker than) expected from a normal-copula null hypothesis.\label{fig_cc2000_nonparam}}
\end{center}
\end{figure}

@@

knitr::kable(tab_3biv_npa_summary, booktabs = T, escape=F,
             format = "latex",linesep = "", row.names=F,
      caption.short = "Nonparametric asymmetry tests against a normal copula null model with birds, mammals and Cedar Creek data",
      caption = "Nonparametric asymmetry tests against an normal copula null model with birds, mammals and Cedar creek data: for birds and mammals data, results should be interpreted as of soil C and N results (see Table \\ref{MT-tab_soilCN_asym} which shows stronger upper tail dependence). For Cedar Creek data, first slice ($FS$) and last slice ($LS$) stand for $(0,0.25)$ and $(0.75,1)$ range of lower and upper bounds. Though the non-parametric result does not indicate significantly stronger lower tail dependence for Cedar Creek data, the 
greater than (>) sign before the numbers do show the opposite direction in tail-dependence compared to other bivariate datasets.\\label{tab_3biv_npa_summary}")
```

@@

For first three bivariate datasets in Table \ref{MT-tab_bivar_summary},
the statistic $P_{0,0.1}-P_{0.9,1}$ was computed for real data and 
for 1000 Kendall-preserving surrogate datasets with normal copulas. Due to relatively smaller sample size,
for Cedar Creek data we used $P_{0,0.25}-P_{0.75,1}$.
For each dataset, the rank of the statistic for real-data in 
the distribution of surrogate values was computed. 
If ranks were $>975$ or $<25$, asymmetry of tail 
dependence was significant compared to a normal copula null hypothesis. Ranks were `r unname(tab_bivar_summary[10,])`
for, respectively, the soil C and N dataset, the bird body masses and metabolic rates dataset, 
the mammal body masses and metabolic rates dataset, and the Cedar Creek data for the year 2000. 
Thus the first three datasets showed stronger 
upper- than lower-tail dependence, and the Cedar Creek data showed the reverse 
(though not significantly), confirming the 
results based on model selection presented in the main text (see 
Table \ref{MT-tab_bivar_summary}, row 9). Model selection results for Cedar Creek showed 
only slightly stronger lower- than upper-tail dependence, consistent with the non-significant 
tendency toward lower-tail dependence also revealed by the non-parametric analysis for that
dataset.

@@

The scheme is identical to that proposed by @bjornstadfalck2001 for application in their now 
widely used spline correlogram method; the same scheme was also used by @walteretal2017 
as part of a method for 
performing model selection among matrix regression models. For each pair of 
locations $i \neq j$, the statistic
of interest, here denoted $s_{ij}$ ($\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, 
$\Dsq_{l_b,u_b}$, $\cor_l-\cor_u$, 
$\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$) was computed for those locations. The 
values $s_{ij}$ were then arranged in
a matrix, $S$, with diagonal entries NA. The mean of the non-NA entries of $S$ was computed. For 
$\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, or $\Dsq_{l_b,u_b}$ this mean characterizes the 
average strength of 
dependence, across all pairs of distinct locations, for data in the portion of the distributions 
given by the bounds $l_b$ and $u_b$. For $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$, 
the mean characterizes average
asymmetry of tail dependence across all pairs of locations. 
Confidence intervals for the value of $\mean(S)$ were then computed as quantiles of a 
distribution of bootstrapped values. If $\mean(S)$ was significantly different from 0, as
revealed by the confidence intervals, for the statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, 
or $\Dsq_u-\Dsq_l$, then tail dependence was significantly asymmetric, in contrast to a 
normal copula null hypothesis.
Bootstrapping was performed by resampling the locations of measurement, with replacement. 
This corresponds to 
resampling, with replacement, the rows and columns of $S$, and is described in detail elsewhere 
[@bjornstadfalck2001 ; @walteretal2017].

@@

tab_asym<-readRDS("./Results/asym_sens_results/stat_npa_asym_b_0.1_b_0.5.RDS")
tab_stat_npa_asym<-c()
for(i in c(1:8)){
  temp<-make_table_stat_fn(data=tab_asym[[i]])
  tab_stat_npa_asym<-rbind(tab_stat_npa_asym,temp)
}
gtext<-c("b=0.1, g$_{1}$","b=0.1, g$_{2}$","b=-0.1, g$_{1}$","b=-0.1, g$_{2}$","b=0.5, g$_{1}$","b=0.5, g$_{2}$","b=-0.5, g$_{1}$","b=-0.5, g$_{2}$")
tab_stat_npa_asym<-cbind(b=rep(c(0.1,0.1,-0.1,-0.1,0.5,0.5,-0.5,-0.5),each=3),
  g=rep(c("g$_{1}$","g$_{2}$","g$_{1}$","g$_{2}$","g$_{1}$","g$_{2}$","g$_{1}$","g$_{2}$"),each=3),tab_stat_npa_asym)
knitr::kable(tab_stat_npa_asym, booktabs = T, longtable=T,
             format = "latex",linesep = "",escape=F,
             caption.short = "Asymmetric sensitivity model, nonparametric statistics results",
             caption ="Asymmetric sensitivity model, nonparametric statistics results. 
The output of simulations from the first model detailed in 
the methods section for Q2 (i.e., Figs \\ref{MT-fig_ms_asym_b_0.1} and 
\\ref{fig_ms_asym_b_-0.1}-\\ref{fig_ms_asym_b_-0.5}) were subjected to nonparamtric statistical 
analyses decribed in the methods section for Q1. The listed statistic was computed for the
output of the model with the indicated b and g, and the value was compared to 1000 values
of the same statistic computed on surrogate time series randomized to have normal copula structure
(see methods for Q1). A table entry $<X$ (respectively, $>X$) indicates the value of the 
given statistic on the data was less (respectively, more) than its value on $X$ of the 
surrogates, so entries of the form $<X$ (respectively, $>X$) for $X$ equal to $975$ or above 
indicate that upper-tail (respectively, lower-tail) dependence was significantly stronger than 
lower-tail dependence compared to a normal copula null hypothesis. `FS' stands for `first slice'
and refers to the bounds 0 to 0.1; `LS' stands for `last slice' and refers to the bounds
0.9 to 1.\\label{tab_stat_npa_asym}")
```

>>>>

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/BMR_results/birds/birds_numbin_10_bivfunctionplot.pdf}
\caption[Nonparametric tests against a normal-copula null, bird body mass and BMR data]{Nonparametric tests for tail dependence and other deviations from normal copula structure, for bird body mass and BMR data. Same format as Fig. \ref{MT-fig_soilCN_nonparam}, see caption of that figure. \label{fig_bmr_birds_nonparam}}
\end{center}
\end{figure}

@@

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/BMR_results/mammals/mammals_numbin_10_bivfunctionplot.pdf}
\caption[Nonparametric tests against a normal-copula null, mammal body mass and BMR data]{Nonparametric tests for tail dependence and other deviations from normal copula structure, for mammal body mass and BMR data. Same format as Fig. \ref{MT-fig_soilCN_nonparam}, see caption of that figure. \label{fig_bmr_mammals_nonparam}}
\end{center}
\end{figure}

@@

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/Ceder_creek_results/2000cc_numbin_4_bivfunctionplot.pdf}
\caption[Nonparametric tests against a normal-copula null, Cedar Creek data]{Nonparametric tests for tail dependence and other deviations from normal copula structure, for Cedar Creek data, year 2000. As described in the main text, the values of the statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$ and $\Dsq_{l_b,u_b}$ for real data (red crosses) were compared to distributions of their values on 1000 Kendall- or Spearman-preserving normal surrogates of the data (blue and green x's show 0.025 and 0.975 quantiles), separately in two comparisons for each of the ranges $(l_b,u_b)=(0,0.25),\ldots,(0.75,1)$. If the red cross was outside the range given by the x's, text at the top of panels indicates the number of surrogate values the real-data value was greater than or less than. For instance, a value $>N$ (respectively, $<N$) means the value of the statistic on real data was greater than (respectively, less than) its value on $N$ surrogates. When the statistic was greater than 975 or less than 975 surrogate values, it indicates significance ($95\%$ confidence level). When $\cor$ or $\Ps$ values (respectively, $\Dsq$ values) were greater than surrogates, it means dependence in that part of the distributions was stronger than (respectively, weaker than) expected from a normal-copula null hypothesis.\label{fig_cc2000_nonparam}}
\end{center}
\end{figure}

@@

knitr::kable(tab_3biv_npa_summary, booktabs = T, escape=F,
             format = "latex",linesep = "", row.names=F,
      caption.short = "Nonparametric asymmetry tests against a normal-copula null model with birds, mammals and Cedar Creek data",
      caption = "Nonparametric asymmetry tests against a normal-copula null model with birds, mammals and Cedar creek data: for birds and mammals data, results should be interpreted as of soil C and N results (see Table \\ref{MT-tab_soilCN_asym} which shows stronger upper tail dependence). For Cedar Creek data, first slice ($FS$) and last slice ($LS$) stand for $(0,0.25)$ and $(0.75,1)$ range of lower and upper bounds. Though the non-parametric result does not indicate significantly stronger lower tail dependence for Cedar Creek data, the 
greater than (>) sign before the numbers do show the opposite direction in tail-dependence compared to other bivariate datasets.\\label{tab_3biv_npa_summary}")
```

@@

For first three bivariate datasets in Table \ref{MT-tab_bivar_summary},
the statistic $P_{0,0.1}-P_{0.9,1}$ was computed for real data and 
for 1000 Kendall-preserving surrogate datasets with normal copulas. Due to relatively smaller sample size,
for Cedar Creek data we used $P_{0,0.25}-P_{0.75,1}$.
For each dataset, the rank of the statistic for real-data in 
the distribution of surrogate values was computed. 
If ranks were $>975$ or $<25$, asymmetry of tail 
dependence was significant compared to a normal-copula null hypothesis. Ranks were `r unname(tab_bivar_summary[10,])`
for, respectively, the soil C and N dataset, the bird body masses and metabolic rates dataset, 
the mammal body masses and metabolic rates dataset, and the Cedar Creek data for the year 2000. 
Thus the first three datasets showed stronger 
upper- than lower-tail dependence, and the Cedar Creek data showed the reverse 
(though not significantly), confirming the 
results based on model selection presented in the main text (see 
Table \ref{MT-tab_bivar_summary}, row 9). Model selection results for Cedar Creek showed 
only slightly stronger lower- than upper-tail dependence, consistent with the non-significant 
tendency toward lower-tail dependence also revealed by the non-parametric analysis for that
dataset.

@@

The scheme is identical to that proposed by @bjornstadfalck2001 for application in their now 
widely used spline correlogram method; the same scheme was also used by @walteretal2017 
as part of a method for 
performing model selection among matrix regression models. For each pair of 
locations $i \neq j$, the statistic
of interest, here denoted $s_{ij}$ ($\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, 
$\Dsq_{l_b,u_b}$, $\cor_l-\cor_u$, 
$\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$) was computed for those locations. The 
values $s_{ij}$ were then arranged in
a matrix, $S$, with diagonal entries NA. The mean of the non-NA entries of $S$ was computed. For 
$\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, or $\Dsq_{l_b,u_b}$ this mean characterizes the 
average strength of 
dependence, across all pairs of distinct locations, for data in the portion of the distributions 
given by the bounds $l_b$ and $u_b$. For $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$, 
the mean characterizes average
asymmetry of tail dependence across all pairs of locations. 
Confidence intervals for the value of $\mean(S)$ were then computed as quantiles of a 
distribution of bootstrapped values. If $\mean(S)$ was significantly different from 0, as
revealed by the confidence intervals, for the statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, 
or $\Dsq_u-\Dsq_l$, then tail dependence was significantly asymmetric, in contrast to a 
normal-copula null hypothesis.
Bootstrapping was performed by resampling the locations of measurement, with replacement. 
This corresponds to 
resampling, with replacement, the rows and columns of $S$, and is described in detail elsewhere 
[@bjornstadfalck2001 ; @walteretal2017].

@@

tab_asym<-readRDS("./Results/asym_sens_results/stat_npa_asym_b_0.1_b_0.5.RDS")
tab_stat_npa_asym<-c()
for(i in c(1:8)){
  temp<-make_table_stat_fn(data=tab_asym[[i]])
  tab_stat_npa_asym<-rbind(tab_stat_npa_asym,temp)
}
gtext<-c("b=0.1, g$_{1}$","b=0.1, g$_{2}$","b=-0.1, g$_{1}$","b=-0.1, g$_{2}$","b=0.5, g$_{1}$","b=0.5, g$_{2}$","b=-0.5, g$_{1}$","b=-0.5, g$_{2}$")
tab_stat_npa_asym<-cbind(b=rep(c(0.1,0.1,-0.1,-0.1,0.5,0.5,-0.5,-0.5),each=3),
  g=rep(c("g$_{1}$","g$_{2}$","g$_{1}$","g$_{2}$","g$_{1}$","g$_{2}$","g$_{1}$","g$_{2}$"),each=3),tab_stat_npa_asym)
knitr::kable(tab_stat_npa_asym, booktabs = T, longtable=T,
             format = "latex",linesep = "",escape=F,
             caption.short = "Asymmetric sensitivity model, nonparametric statistics results",
             caption ="Asymmetric sensitivity model, nonparametric statistics results. 
The output of simulations from the first model detailed in 
the methods section for Q2 (i.e., Figs \\ref{MT-fig_ms_asym_b_0.1} and 
\\ref{fig_ms_asym_b_-0.1}-\\ref{fig_ms_asym_b_-0.5}) were subjected to nonparamtric statistical 
analyses decribed in the methods section for Q1. The listed statistic was computed for the
output of the model with the indicated b and g, and the value was compared to 1000 values
of the same statistic computed on surrogate time series randomized to have normal copula structure
(see methods for Q1). A table entry $<X$ (respectively, $>X$) indicates the value of the 
given statistic on the data was less (respectively, more) than its value on $X$ of the 
surrogates, so entries of the form $<X$ (respectively, $>X$) for $X$ equal to $975$ or above 
indicate that upper-tail (respectively, lower-tail) dependence was significantly stronger than 
lower-tail dependence compared to a normal-copula null hypothesis. `FS' stands for `first slice'
and refers to the bounds 0 to 0.1; `LS' stands for `last slice' and refers to the bounds
0.9 to 1.\\label{tab_stat_npa_asym}")
```
